\documentclass{beamer}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amssymb}
\newcommand{\n}{\sim}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\ceil}[1]{\left\lceil #1 \right\rceil}

\title{MATH1061 Week 2 Tutorial}
\subtitle{"Do you want to go to the maths tutorial or skip it?"
"Yes."}
\date{5 March 2013}

\begin{document}

\frame{\titlepage}

\begin{frame}
\frametitle{Hello!}

Things you can do in tutorials:
\pause
\begin{itemize}
\item One more chance to understand last week's lectures.
\pause
\item Get practice and tips in answering questions.
\pause
\item Discuss tasks with others (e.g. assignment 1 Q7).
\pause
\item Get individual help on assignment questions.
\pause
\item Get feedback on marked assignments.
\end{itemize}
\pause

Maths First Year Tutuoring: To get individual help, go to
Priestley-642 from 2-4pm any weekday.
\pause

Tutorial questions will come from \{Assignments,
"Exercises" boxes in workbook, my brain\}.


\end{frame}

\begin{frame}
\frametitle{Logical Statments}

What is the logical negation of each of these statements?
Which is true - the original statement or its negation? (Check by seeing that
the other one is false)
\pause

\begin{enumerate}
\item Our tutorial is in the Priestley building, or you have to submit
your assignments in the Priestley building.
\item If pigs can fly, then Mr. Burns will donate a million dollars to the
local orphanage.\footnote{For this exercise, it is safe to assume that pigs
can't fly. For Mr. Burns, it's not a safe assumption.}
\item If there are Engineering students in the Hawken building, then there will
be a Discrete Maths lecture on Wednesday morning.
\item If the sky was pink, then the sky would be blue.
\end{enumerate}

\end{frame}

\begin{frame}
\frametitle{Conditionals}

The following statement is true: "Jackson brings his umbrella to uni when it's
going to rain". \pause Which of the following conclusions can we make?

\begin{enumerate}
\item To rain, it suffices that Jackson has his umbrella at uni.
\item If it is raining, then Jackson brought his umbrella.
\item If Jackson doesn't have his umbrella, then it's not raining.
\item If it's not going to rain, Jackson didn't bring his umbrella.
\item It's necessary that Jackson has his umbrella for it to rain.
\item A rainy forecast causes Jackson to bring his umbrella to uni.
\end{enumerate}

\end{frame}

\end{document}
